### 2013-07-09 MHCslides/MHCslides.tex
started mhc slides

% -*- TeX -*- -*- UK -*- -*- BMR -*-
% ----------------------------------------------------------------
% Beamer presentation ************************************************
% **** -----------------------------------------------------------
\documentclass{beamer}%[hyperref={backref=slide}]

\input{sl_slide_preamble.tex}
%\input{sl_slide_preamble_nonotes.tex}
\input{sl_slide_graphics_preamble.tex}
\input{sl_definitions.tex}
\input{sl_slide_symbols.tex}
%
\graphicspath{{"Figures/"}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\I}{\mathbf{I}}
%prob vector
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\wv}{\vec{w}}
%super/subscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
%quantities
\newcommand{\initial}{\mathcal{I}}
\newcommand{\area}{\mathcal{A}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\comp}{^\mathrm{c}}
\renewcommand{\e}{\mathsf{e}}
%---------Title-----------------------------------------------------------

\title[Impaired learning with enhanced plasticity]{A saturation model for impaired learning with enhanced plasticity}
%
\subtitle{\small{based on work in preparation by: T.D. Barbara Nguyen-Vu, Grace Q. Zhao, Han-Mi Lee, SL, Surya Ganguli, Carla J. Shatz, Jennifer L. Raymond
}}
%
\author{Subhaneil Lahiri%\inst{1}
}
%
\institute[Stanford]{%
%\inst{1}
Stanford University, Applied Physics
}
%
%\slideCaption{}

%---------Beginning--------------------------------------------------------

\begin{document}

%-------------Slide--------------------------------------------------------

\begin{frame}
%
 \titlepage
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Outline}
%
 \tableofcontents[hideallsubsections]
 %
 \note[item]{review terrible properties of simple synapses.}
 \note[item]{mathematical formalism of model, quantify performance (memory decay over time)}
 \note[item]{upper bounds on single numbers that depend on whole memory curve (decay over time)}
 \note[item]{upper bounds at finite times}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}[allowframebreaks]{References}
%

 {\small
 \bibliographystyle{unsrt_slides}
 \bibliography{maths,neuro}
 }
%
\end{frame}

%-----End----------------------------------------------------------------

\end{document}