### 2013-07-12 MHCslides/MHCslides.tex
expt results section

% -*- TeX -*- -*- UK -*- -*- BMR -*-
% ----------------------------------------------------------------
% Beamer presentation ************************************************
% **** -----------------------------------------------------------
\documentclass{beamer}%[hyperref={backref=slide}]

\usepackage{array}
\usepackage{cancel}
\input{sl_slide_preamble.tex}
%\usecolortheme[named=darkblue]{structure}
\usecolortheme[rgb={0.6372549 0.27843137 0.27058824}]{structure}
%\setbeamercolor{alerted text}{fg=blue!80!black}
%\input{sl_slide_preamble_nonotes.tex}
\input{sl_slide_graphics_preamble.tex}
\input{sl_definitions.tex}
\input{sl_slide_symbols.tex}
%
\graphicspath{{"Figs/"}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\I}{\mathbf{I}}
%prob vector
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\wv}{\vec{w}}
%super/subscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
%quantities
\newcommand{\initial}{\mathcal{I}}
\newcommand{\area}{\mathcal{A}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\comp}{^\mathrm{c}}
\renewcommand{\e}{\mathsf{e}}
%---------Title-----------------------------------------------------------

\title[Saturation by enh. plasticity impairs learning]{A saturation model for impaired learning with enhanced plasticity}
%
\subtitle{\small{based on work in preparation by: T.D. Barbara Nguyen-Vu, Grace Q. Zhao, Han-Mi Lee, SL, Surya Ganguli, Carla J. Shatz, Jennifer L. Raymond
}}
%
\author{Subhaneil Lahiri%\inst{1}
}
%
\institute[Stanford]{%
%\inst{1}
Stanford University, Applied Physics
}
%
%\slideCaption{}

%---------Beginning--------------------------------------------------------

\begin{document}

%-------------Slide--------------------------------------------------------

\begin{frame}
%
 \titlepage
 \note[item]{Acknowledge Barbara and Grace}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Introduction}
%
 One might expect enhanced plasticity to enhance learning.
 \note[item]{It does help in some cases}
 But often it actually causes impairment.

 \vp We argue that this impaired learning is due to basal activity biasing the synapses prior to learning, leaving fewer synapses available for further plasticity.

 \vp By analysing several models of complex synapses, we see that the behaviour of enhanced LTD mice in a motor learning task can constrain the synaptic structure.
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Outline}
%
 \tableofcontents[hideallsubsections]
 %
%
\end{frame}

%-------------Section--------------------------------------------------------

\section{VOR learning and the cerebellum}

%-------------Slide--------------------------------------------------------

\begin{frame}{Vestibular Occular Reflex}
%
 \parbox[t]{0.4\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VOR.svg}}}
 \parbox[t]{0.59\linewidth}{%
 Eye movements compensate for head movements to maintain fixation.

 \vp Requires control of $\text{VOR gain} = \frac{\text{eye velocity}}{\text{head velocity}}$.

 \vp Needs to be adjusted as eye muscles age, \etc
 }

 \citerr{duLac1995review,Boyden2004review}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{VOR training}
%
 \parbox[t]{0.2\linewidth}{%
   \aligntop{\includegraphics[width=0.99\linewidth]{VORinc.svg}}\\[1cm]
     \note[item]{trick brain into thinking VOR gain needs adjusting my moving visual stimulus}
     \note[item]{anti-phase \lto increase gain}
   \aligntop{\includegraphics[width=0.99\linewidth]{VORdec.svg}}
     \note[item]{in phase \lto decrease gain}
   }
 \hspace{0.15\linewidth}
 \parbox[t]{0.59\linewidth}{%
   \aligntop{\includegraphics[width=0.99\linewidth]{VORcircuit.svg}}

   \vp PF-Pk: \parbox[t]{0.5\linewidth}{PF+CF \lto LTD, \\
              PF+\cancel{CF} \lto LTP.}
 }
   \note[item]{Gain change involves cerebellum}
   \note[item]{Marr-Albus-Ito: Pf-Pk synapses}
   \note[item]{Lisberger-Miles: Vestibular input-VN synapses}
   \note[item]{Different mechs for different freq, head angle, gain up/down.}
   \note[item]{Different Pk cells have different tunings.}
   \note[item]{Gain up in case of interest: LTD in Pf-Pk in flocculus}
   \note[item]{Gain down: uses different mech for behaviour, but does reverse LTD in Pf-Pk in flocculus}

   \citerr{Marr1969cerebellum,Albus1971cerebellum,ito1972cerebellum,Ito1982VOR,Miles1981VOR}
%
\end{frame}

%-------------Section--------------------------------------------------------

\section{The effects of enhanced plasticity and saturation}

%-------------Slide--------------------------------------------------------

\begin{frame}{MHC-I D$^\mathrm{b}$K$^\mathrm{b} -/-$ knockout}
%
 Knockout of molecule lowers threshold for LTD in PF-Pk synapses.\\
 \citerr{McConnell2009MHCcerebellum}
 \note[item]{Easier LTD \lto expect better learning}

 \vp \begin{center}
 \begin{tabular}{ccc}
   Baseline & \hspace{0.1\linewidth} & Gain increase at $1.0\,\mathrm{Hz}$ \\
   \alignmid{\includegraphics[width=0.4\linewidth]{baseline.svg}} && 
   \alignmid{\includegraphics[width=0.3\linewidth]{gain_inc.svg}} 
 \end{tabular}
 \end{center}
 \note[item]{No difference at baseline}
 \note[item]{Impairment of learning}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Saturation hypothesis}
%
 \begin{center}
 \begin{tabular}{ccc}
   Error model & \hspace{0.1\linewidth} & Saturation model \\[1cm]
   \alignmid{\includegraphics[width=0.3\linewidth]{err_model.svg}} &&
   \alignmid{\includegraphics[width=0.3\linewidth]{sat_model.svg}}
 \end{tabular}
 \end{center}
 \note[item]{Older explanation: error model}
 \note[item]{Our model: baseline activity \lto saturation \lto less depression possible}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Evidence: level of depression}
%
 \aligntop{\includegraphics[width=0.3\linewidth]{pSer880.svg}}
 \hp
 \parbox[t]{0.6\linewidth}{%
 Basal level of GluR2 phosphorylation at serine 880 in AMPA receptor.

 \vp Biochemical signature of PF-Pk LTD.

 \vp Shows that \# depressed synapses in flocculus is larger in KO than WT.
 \note[item]{Predicted by saturation hypothesis}
 }
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Evidence: saturation by CF stimulation}
%
 Use Channelrhodopsin to stimulate CF \parbox[t]{0.45\linewidth}{%
 \lto increase LTD in PF-Pk synapses \\
 \lto simulate saturation in WT.
 }
 \note[item]{should result in similar behaviour to KO}

 \vp\begin{center}
   \alignmid{\includegraphics[width=0.4\linewidth]{CFstim.svg}}
   \hspace{0.1\linewidth}
   \alignmid{\includegraphics[width=0.4\linewidth]{CFstim_results.svg}}
 \end{center}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Evidence: desaturation by gain-decrease pretraining}
%
 %
 \begin{center}
   \includegraphics[width=0.7\linewidth]{desat.svg}

   \vp \includegraphics[width=0.7\linewidth]{gain_dec.svg}
 \end{center}
 %
 \note[item]{gain dec reverses LTD}
 \note[item]{but behaviour from elsewhere}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Summary of training results}
%
 \begin{center}
   \alignmid{\includegraphics[width=0.6\linewidth]{data_shifted.svg}}
   \hspace{0.05\linewidth}
   \alignmid{\includegraphics[width=0.3\linewidth]{comparisons.svg}}
 \end{center}
 \note[item]{Restricted to gain inc for comparison}
 \note[item]{Black: WT. Red: KO}
 \note[item]{Solid: no pre. Dashed: with pre}
 \note[item]{Horz and vert comparisons: conceptual}
 \note[item]{Diagonal comparisons: paramter fitting. Depend on size of KO vs.\ pretraining}
 \note[item]{KO hurts w/o, but helps w/}
 \note[item]{pre helps KO but hurts WT}
 \note[item]{top and left most restrictive}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Questions}
%
 \begin{itemize}
   \item Can the saturation effect overcome the enhanced plasticity?
   \note[item]{in competition}
   \item How can a little reverse bias help, but too much hurt?
   \note[item]{first makes sense, but second?}
   \item Can we find a purely synaptic explanation of these results?
   \note[item]{This is a question about synaptic populations after all.}
 \end{itemize}
%
\end{frame}

%-------------Section--------------------------------------------------------

\section{Modelling approach}

%-------------Slide--------------------------------------------------------

\begin{frame}{Model of circuit}
%
 %
 \begin{center}
   \includegraphics[width=0.9\linewidth]{VORmodel.svg}
 \end{center}
 %
 \note[item]{Contralateral baseline shift compensates for Our baseline shift}
 \note[item]{Gain increase due to LTD at lightning}
 \note[item]{Gain decease due to plasticity elsewhere, but also reverses LTD at lightning.}
 \note[item]{Nonlinearity here won't affect our questions, as long as it doesn't change}
 \note[item]{Nonlinearity before compensation could change things}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Complex synapses}
%
  \aligntop{\includegraphics[width=5cm]{synapse.svg}}\hp
  %
  \note[item]{functional states, not molecules}
  \note[item]{synaptic weight depends on state}
  \note[item]{many states can have same weight}
  %
  \hp\aligntop{\includegraphics[width=2cm]{pot.svg}}
  \hp\aligntop{\includegraphics[width=2cm]{dep.svg}}
  %
  \note[item]{stochastic transitions}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Simplifying assumptions}
%
\begin{itemize}
  \item No spatial/temporal correlations in plasticity events.
      %
      \note[item]{allows us to concentrate on synapse, not neuron/network}
      %
%  \item Which synapses eligible for plasticity chosen randomly.
      %
      \note[item]{No filing system}
      %
  \item Potentiating/depressing plasticity events $\sim$ Poisson processes.% with rates $rf\potdep$, where $f\pot+f\dep=1$.
      %
      \note[item]{don't care if STDP...}
%      %
%      \note[item]{$r=$ total rate of plasticity events per synapse, $f\potdep=$ fraction of events that are potentiating/depressing.}
      %
  \item Potentiation and depression are described by Markov processes.% with transition probabilities $\M\potdep$.
%      %
%      \note[item]{matrix elements: transition prob from $i\to j$, given pot/dep}
%  \item Ideal observer: read weights directly.
%      \note[item]{not electrical activity: don't model neurons/network}
%      \note[item]{upper bound on electrical activity readout}
%      %
%  \item Synaptic weights can only take values $\pm1$.
%      %
%      \note[item]{looks like binary synapse from outside. Inside...}
      %
  \citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
\end{itemize}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Synaptic dynamics}
%
  There are $N$ identical synapses with $M$ internal functional states.
  %
  \begin{center}
    \includegraphics[width=0.7\linewidth]{model.svg}
  \end{center}
  %
  \note[item]{stoch process has steady state.}
  \note[item]{Prior activity puts it in this state. row vec.}
  \note[item]{plasticity events at rate r}
  \note[item]{fraction pot/dep}
  \note[item]{probs changed by Markov matrices, prob $i\to j$}
  \note[item]{Readout: synaptic weight vec when in each state.}
  \vp
% At $t=0$, the memory is created by $\M\potdep$ with probability $f\potdep$.
% \note[item]{for this one, we keep track of pot/dep, look for inc/dec of $\w$}
%
% \vp Forgetting caused by subsequent memories, evolving as
%
  \begin{equation*}
  \begin{aligned}
    \diff{\pr(t)}{t} &= r\pr(t)\frg,
    &\qquad
    \frg &= f\pot\M\pot+f\dep\M\dep-\I,\\&&
    \eq\frg &=0.
  \end{aligned}
  \end{equation*}
%
  \note[item]{Memory at $t=0$, keep track of pot/dep}
  \note[item]{subsequent: average over pot/dep}
% \note[item]{$\frg$ is forgetting matrix, $\I=$identity, don't keep track of pot/dep}
% Eventually, this will settle into the equilibrium distribution:
% %
% %
% \note[item]{In equilibrium prior to memory creation}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Modelling VOR learning}
%
\begin{tabular}{>{\bfseries}ll>{\impl}l}
    Mutation: &
    Changes mechanism of LTD  &
    change $\M\dep$.
    \note[item]{lower threshold \lto increase off-diagonal elements.}
 \\[1cm]
    Training: &
    Changes statistics of LTP/LTD  &
    change $r,f\pot,f\dep$.
    \note[item]{Only parameters we have. Don't care about $r$.}
    \note[item]{Same PF+CF input \lto same $r,f\pot,f\dep$ in each case.}
 \\[1cm]
    Learning: &
    Change in VOR gain &
    fn.\ of decrease in $\av{\w}$.
    \note[item]{Only output we have. Don't keep track of synaptic identity.}
    \note[item]{Input to Pk, some linear combination of $\w$'s. }
\end{tabular}
%
\end{frame}

%-------------Section--------------------------------------------------------

\section{Modelling results}

%-------------Slide--------------------------------------------------------

\begin{frame}{Binary synapse}
%
 %
 \begin{center}
   \includegraphics[width=0.15\linewidth]{binary.svg} \\[1cm]
   \aligntop{\includegraphics[width=0.35\linewidth]{binary_learnS.eps}}
   \hspace{2cm}
   \aligntop{\includegraphics[width=0.35\linewidth]{comparisons_binary.svg}}
 \end{center}
 %
 \note[item]{Compare solid curves}
 \note[item]{Compare black curves}
 \note[item]{understand why next slide}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Binary synapse: initial distributions}
%
 %
\adjustbox{valign=m}{\parbox[t]{0.4\linewidth}{
  \begin{raggedleft}
  Learning rate: \phantom{rate of events} \\
  \hp\hp\parbox{0.75\linewidth}{%
  $\sim$ rate of events\\
  $\times$ prob.\ of transition\\
  $\times$ prob.\ ready for $\Delta w$\\
  $\times$ (-$\Delta w$)}
  \end{raggedleft}
}}
% \begin{center}
\alignmid{\parbox[t]{0.5\linewidth}{
 \begin{tabular}{ccc}
   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_wt_wo.svg}}
   & \alignmid{\includegraphics[height=0.12\linewidth]{left_bad.svg}} &
   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_ko_wo.svg}}
   \\[1.5cm] \alignmid{\includegraphics[height=0.12\linewidth]{up_bad.svg}}
   && \alignmid{\includegraphics[height=0.12\linewidth]{up_good.svg}}\\[0.5cm]
   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_wt_w.svg}}
   & \alignmid{\includegraphics[height=0.12\linewidth]{left_good.svg}} &
   \alignmid{\includegraphics[height=0.5\linewidth]{binary_bar_ko_w.svg}}
 \end{tabular}
}}
% \end{center}
 %
 \note[item]{WT: start with everything equal -- just for illustration, not essential}
 \note[item]{WT: during training, increase $f\dep$ (green arrow) $\to$ weakening.}
 \note[item]{KO: inc $q\pot\to$ bias}
 \note[item]{KO: competition between inc prob trans \& dec prob ready}
 \note[item]{KO: first one wins. see why after next model}
 \note[item]{pre: reduces/reverses bias. always helps.}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Serial synapse}
%
 %
 \begin{center}
   \includegraphics[width=0.5\linewidth]{serial.svg}\\
   \citerr{Leibold2008serial} \\[1cm]
 \note[item]{Still looks binary from outside. Hidden states (not essential).}
 \note[item]{Only see $\Delta w$ at boundary.}
   \aligntop{\includegraphics[width=0.35\linewidth]{serial_learnS.eps}}
   \hspace{2cm}
   \aligntop{\includegraphics[width=0.35\linewidth]{comparisons_serial.svg}}
 \end{center}
 %

 \note[item]{understand why next slide}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Serial synapse: initial distributions}
%
 %
 \begin{center}
 \begin{tabular}{ccc}
   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_wt_wo.svg}}
   & \alignmid{\includegraphics[height=0.05\linewidth]{right_good.svg}} &
   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_ko_wo.svg}}
   \\[1.5cm] \alignmid{\includegraphics[height=0.05\linewidth]{down_good.svg}}
   && \alignmid{\includegraphics[height=0.05\linewidth]{up_good.svg}}\\[0.5cm]
   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_wt_w.svg}}
   & \alignmid{\includegraphics[height=0.05\linewidth]{left_good.svg}} &
   \alignmid{\includegraphics[height=0.19\linewidth]{serial_bar_ko_w.svg}}
 \end{tabular}
 \end{center}
 %
 \note[item]{WT: start with everything equal -- just for illustration, not essential}
 \note[item]{WT: during training, increase $f\dep$ (green arrow) $\to$ weakening.}
 \note[item]{KO: inc $q\pot\to$ bias, now exponential}
 \note[item]{KO: comp. between inc prob trans \& dec prob ready, now only at bndry}
 \note[item]{KO: second one wins, now exponential}
 \note[item]{pre: reduces/reverses bias.}
 \note[item]{pre: little reverse bias repopulates bndry, helps.}
 \note[item]{pre: too much reverse bias moves away from bndry, hurts.}
 \note[item]{maths next slide}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Mathematical explanation}
%
 Serial synapse: $\eq_i \sim \CN \prn{\frac{q\pot}{q\dep}}^i$.

 \vp Learning rate $\sim \eq_{M/2}\prn{\frac{q\dep}{q\pot}} = \CN \prn{\frac{q\pot}{q\dep}}^{\frac{M}{2}-1}$.
 \note[item]{Detailed balance. Exponential decay.}

 \vp For $M>2$: larger $q\dep \implies$ slower learning.
 \note[item]{for large enough $M,q\pot$, overcome $\CN$}

 \vp For $M=2$: larger $q\dep \implies$ larger $\CN \implies$ faster learning.
 \note[item]{Other factor in $\eq$ smaller $\implies \CN$ larger.}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Essential features}
%
 The success of the serial model relies on two features:
 \begin{itemize}
   \item Enhancing the effect of saturation,
   \note[item]{due to exponential decay}
   \item Metaplasticity -- repeated potentiation makes subsequent depression harder.
   \note[item]{push away from boundary where signal generated}
 \end{itemize}
 %
 \begin{center}
   \includegraphics[width=0.4\linewidth]{serial.svg}
 \end{center}
 %
 \note[item]{borne out by other models that fail/succeed}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Other models that fail}
%
 \begin{tabular}{ll}
   % & for next tab, \\ for new line...
    \aligntop{\includegraphics[width=0.4\linewidth]{multistate.svg}}\hspace{1.5cm} &
    \aligntop{\includegraphics[width=0.28\linewidth]{multistate_learnS.eps}}
    \note[item]{MS: linear weights, unlike serial.}
    \note[item]{like bunch of binary synapses in series.}
    \note[item]{solid curves: fails early on , but catches up quickly}
    \note[item]{black curves: fails badly}
    \note[item]{No real enhancement of saturation, no metaplasticity.}
    \note[item]{All transitions contribute: pushing to end has little effect.}
    \\[2cm]
    \aligntop{\includegraphics[width=0.4\linewidth]{pooled.svg}} &
    \aligntop{\includegraphics[width=0.28\linewidth]{pooled_scarce_learnS.eps}}
    \note[item]{Pooled: resource depleted by pot/dep. replenished by reverse.}
    \note[item]{solid curves succeed: enhanced saturation}
    \note[item]{black curves fail: opposite metaplasticity, pot makes dep easier}
 \end{tabular}

 \citerr{amit1994learning}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Other models that work}
%
 \begin{tabular}{ll}
   % & for next tab, \\ for new line...
    \aligntop{\includegraphics[width=0.4\linewidth]{multistate_nonuni.svg}}\hspace{1.5cm} &
    \aligntop{\includegraphics[width=0.28\linewidth]{nonuni_learnS.eps}} \\[2cm]
    \aligntop{\includegraphics[width=0.4\linewidth]{cascade.svg}} &
    \aligntop{\includegraphics[width=0.28\linewidth]{cascade_long_learnS.eps}}
 \end{tabular}
 \note[item]{Both models, trans probs decay exponentially from centre.}
 \note[item]{Nonuni: linear weights. Cascade: binary weights.}
 \note[item]{Enhanced saturation and metaplasticity}
 \note[item]{Pushing to end makes pot and dep harder}
 \note[item]{Note: hidden states not necessary}

 \citerr{Fusi2005cascade}
%
\end{frame}

%-------------Slide--------------------------------------------------------

\begin{frame}{Conclusions and further questions}
%
 \begin{itemize}
   \item The saturation effect overcome the enhanced plasticity, if it is enhanced.
   \alert{Requires complexity}
   \note[item]{\eg exponential deacy, resource depletion,\ldots}
   \item A little reverse bias can help, but too much hurts, if repeated potentiation makes depression harder.
   \alert{Requires metaplasticity}
   \note[item]{\eg moving away from weight boundary, or weaker transitions.}
   \item We can find a purely synaptic explanation of VOR behaviour, iff the synapses have these features.
   \note[item]{Other explanations? Non-linearity in PK cell?}
   \item We used behaviour to constrain molecular structure of synapses!
   \item Can we constrain it further with more experiments?
 \end{itemize}
%
\end{frame}

%%-------------Slide--------------------------------------------------------
%
%\begin{frame}{Acknowledgements}
%%
% 
%%
%\end{frame}
%
%-------------Slide--------------------------------------------------------

\begin{frame}[allowframebreaks]{References}
%

 {\small
 \bibliographystyle{unsrt_slides}
 \bibliography{maths,neuro}
 }
%
\end{frame}

%-----End----------------------------------------------------------------

\end{document}