### 2013-08-02 NIPS/rebuttal.txt
working on NIPS rebuttal

Reviewer 1:
Regarding the discrepancy between the envelope and the numerical results: the envelope is just an upper bound and we 
do not claim that it is a tight bound. The reason is that equation (18) is not aa complete set of constraints, as 
discussed below it. In the paragraph from line 377 to 400, we discuss this point further. Note that the numerical 
results are not always instructive as the numerical procedures can be prevented from reaching the global maximum by 
local maxima. This is shown by the fact that our hand designed models can outperform the numerical methods at late 
times. In fact, the apparent drop-off of the solid red curve at very late times comes from not allowing small enough 
epsilon. This should be fixed in fig5.

Reviewer 2:
We'll add a derivation of equation 3 to the supplement. We could add a description in the main text along
the lines of: "The factor of p^infinity describes the synapses being in the steady-state distribution before the
memory is encoded. The factor of (M^pot-M^dep) comes from the encoding of the memory at t=0, with w_ideal being +/-1 
in synapses that are potentiated/depotentiated. The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity events and the factor of w indicates the readout 
via the synaptic weight."
I think we referred to all of the original parts of the supplement in the main paper. I'm not sure how we could
incorporate it into the main paper without exceeding the page limit.

Reviewer 3:
line 195: in the introduction, we cited [17], which describes diversity in synaptic structure across the vertebrate 
brain. This could be related to optimisation for different timescales, but anything more than speculation would 
require a better understanding of the relation between structure and function (which is what we intend to begin with 
this work).
line 220:
line 294: agreed.
line 299: The fact that eta_i depends on M^pot/dep is taken into account. In equation (57) of the supplement, the 
term involving c_g comes from this dependence. If the reviewer is concerned that tho order of the eta_i could change 
during the maximisation procedure: note that necessary conditions for a maximum only require that there is no 
infinitesimal perturbation that increases the area. Therefor we need only consider an infinitesimal neighbourhood of 
the model, in which the order will not change.
line 304: Yes. We'll add this phrase to the text.
line 326: In fig2(a,b), we need not have an equal number of w=+/-1 states (we could change the figure to reflect 
this), but those states are not functionally relevant, as shown by their equivalence to fig2c. The model in fig2c, of 
course, has no room for such asymmetry. The models in figs4,5b do need to have equal numbers of +/- states, asymmetry 
would make them worse. For fig4, the effect of the asymmetry would be reduced in the limit as epsilon -> 0, as all
states other than the end states would have very small p^infinity.
line 415: It is only for t_0 > sqrt(N)M that the models that nearly touch the envelope are linear chains. The model 
in fig5b is not a linear chain, as it has shortcut transitions, but they are only the best models (that we have 
found) for times t_0 < sqrt(N)M. Would it help if we repeated fig4 as fig5c?