### 2013-08-02 NIPS/rebuttal.txt
rebuttal comments done

Reviewer 1: We briefly discussed a suggestion for biological
experiments to test the theory in line 428. We could expand on
this. If we measured pre and post-synaptic spike trains, and also
recorded changes in post-synaptic potentials to measure changes in
synaptic weight, we could use hidden Markov model techniques to
find the best-fit synaptic model. However, as shown in this paper,
which model is best depends on the range of time-scales of
interest, so any predictions would also require understanding of
what timescales the particular synapse needs to operate on.
Regarding the discrepancy between the envelope and the
numerical results: the envelope is just an upper bound and we do
not claim that it is a tight bound. The reason is that equation
(18) is not aa complete set of constraints, as discussed below it.
In the paragraph from line 377 to 400, we discuss this point
further. Note that the numerical results are not always instructive
as the numerical procedures can be prevented from reaching the
global maximum by local maxima. This is shown by the fact that our
hand designed models can outperform the numerical methods at late
times. In fact, the apparent drop-off of the solid red curve at
very late times comes from not allowing small enough epsilon. This
should be fixed in fig5.
Yes, the level of noise tolerance is an important consideration.
Note that the stochasticity of the models is an expression of
biological noise. In fact, the type of noise described by the
reviewer could be included by adding extra states to the model with
the different transition rates. However, the synapses should not be
allowed to optimise all of this noise away. It would be interesting
to see how our results change if we put lower limits on the noise
levels. I think this would be beyond the scope of this work, but
should be added to the discussion section.

Reviewer 2: We'll add a derivation of equation 3 to the supplement.
We could add a description in the main text along the lines of:
"The factor of p^infinity describes the synapses being in the
steady-state distribution before the memory is encoded. The factor
of (M^pot-M^dep) comes from the encoding of the memory at t=0, with
w_ideal being +/-1 in synapses that are potentiated/depotentiated.
The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity
events and the factor of w indicates the readout via the synaptic
weight."
I think we referred to all of the original parts of the supplement
in the main paper. I'm not sure how we could incorporate more into
the main paper without exceeding the page limit.

Reviewer 3:
line 195: in the introduction, we cited [17], which describes
diversity in synaptic structure across the vertebrate brain. This
could be related to optimisation for different timescales, but
anything more than speculation would require a better understanding
of the relation between structure and function (which is what we
intend to begin with this work).
line 220: Yes, fig2a,b is an example of why it is difficult to map
molecular states to functional states as, despite appearances,
these models actually only have two functional states due to their
equivalence to fig 2c. One experimental investigation of this could
be along the lines of the experiment described in line 428 (see
first paragraph of our reply to reviewer 1 above for elaboration).
Presumably we will find fewer functional states than molecular
states. Making the link between molecular and functional states
will, unfortunately, be much harder
line 294: agreed.
line 299: The fact that eta_i depends on M^pot/dep is taken into
account. In equation (57) of the supplement, the term involving c_g
comes from this dependence. If the reviewer is concerned that tho
order of the eta_i could change during the maximisation procedure:
note that necessary conditions for a maximum only require that
there is no infinitesimal perturbation that increases the area.
Therefor we need only consider an infinitesimal neighbourhood of
the model, in which the order will not change.
line 304: Yes. We'll add this phrase to the text.
line 326: In fig2(a,b), we need not have an equal number of w=+/-1
states (we could change the figure to reflect this), but those
states are not functionally relevant, as shown by their equivalence
to fig2c. The model in fig2c, of course, has no room for such
asymmetry. The models in figs4,5b do need to have equal numbers of
+/- states, asymmetry would make them worse. For fig4, the effect
of the asymmetry would be reduced in the limit as epsilon -> 0, as
all states other than the end states would have very small
p^infinity.
line 415: It is only for t_0 > sqrt(N)M that the models that nearly
touch the envelope are linear chains. The model in fig5b is not a
linear chain, as it has shortcut transitions, but they are only the
best models (that we have found) for times t_0 < sqrt(N)M. Would it
help if we repeated fig4 as fig5c?