### 2014-07-03 Notes/Identification.tex
synapse fitting notes

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Pib}{\boldsymbol{\Pi}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\pr}{\mathbb{P}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:

\title{Complex Synapse Identification}%
\author{Subhaneil Lahiri}%

% ----------------------------------------------------------------
\begin{document}
% ----------------------------------------------------------------
\maketitle
% ----------------------------------------------------------------
\begin{abstract}
  Notes on using HMM techniques to fit models of complex synapses
\end{abstract}
% ----------------------------------------------------------------

\section{Introduction \label{sec:intro}}

\section{Notation \label{sec:notation}}

Let $t$ be the time of a plasticity event, assumed to be an integer. 
We denote a sequence of event times by $[0,T]$.

We denote states of a synapse by subscripts $i,j$. 
The state at a particular time during a sequence is $i(t)$.
An entire trajectory (sequence of states) is denoted by $i[0,T]$.

The synaptic weight is denoted by $\w$.
Its value when the synapse is in state $i$ is $\w_i$.
Its observed value at time $t$ during a sequence is $\w(t)$.
An entire sequence of observations is denoted by $\w[0,T]$.

The different types of plasticity (\eg potentiation and depression) are indicated by superscripts $\mu,\nu$.
The type of plasticity used at time $t$ is $\mu(t)$.
An entire sequence of plasticity types is denoted by $\mu[0,T]$.

When we are combining data from several sequences of plasticity events, the individual sequences will be denoted by superscripts $a,b$.
The set of all sequences will be denoted by the superscript $A$.

The Markov matrices describing plasticity are $\M^\mu_{ij}$.
The probabilities of the initial states are $\pib_i$.
The set of parameters defining a model (\ie $\brc{\M^\mu_{ij}}$ and $\pib_i$) are denoted by $\M$ for convenience.

\section{Expectation-Maximisation (EM) \label{sec:em}}

When we don't have any prior information about the model $\M$, we fit the model by maximising the likelihood function:
%
\begin{equation}\label{eq:like}
  \pr\cond{\w^A[0,T]}{\M} = \prod_a \pr\cond{\w^a[0,T]}{\M}.
\end{equation}
%
To do this, we define an axillary function
%
\begin{equation}\label{eq:emq}
  Q(\M,\M') = \sum_{i^A[0,T]} \pr\cond{i^A[0,T],\w^A[0,T]}{\M} \log\!\! \brk{
      \frac{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M} } }.
\end{equation}
%
It can be shown that, if $\M'$ is chosen to maximise this quantity (which is easier than maximising the likelihood), the it will have a higher likelihood than $\M'$.
In effect, $\M$ is used to estimate the hidden state trajectories, which are then used to re-estimate the model.
This can be done iteratively to find a local maximum of the likelihood.
If this is repeated several times with random initialisation, we can hope to find the global maximum.

When we have several sequences, we can write
%
\begin{multline}\label{eq:emqseveral}
%\begin{aligned}
  Q(\M,\M') = \sum_{i^A[0,T]} \brk{ \prod_a \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } \brk{\sum_a  \log\!\!  \brk{
      \frac{ \pr\cond{i^A[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^a[0,T]}{\M} } } }\\
   = \sum_a \prod_{b \neq a} \pr\cond{\w^b[0,T]}{\M} \sum_{i^a[0,T]} \pr\cond{i^a[0,T],\w^a[0,T]}{\M} \log\!\!  \brk{
      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
   = \pr\cond{\w^A[0,T]}{\M} \sum_a \frac{Q^a(\M,\M')}{\pr\cond{\w^a[0,T]}{\M}}.
%\end{aligned}
\end{multline}
%

\subsection{Single sequence (Baum-Welch) \label{sec:bw}}

Suppose we have a single sequence of plasticity events.
Then the joint likelihood of a trajectory and sequence observed synaptic weights is given by
%
\begin{equation}\label{eq:trajlike}
\begin{aligned}
  \pr\cond{t[0,T], \w[0,T]}{\M} &= \pib_{i(0)} \, \Pib(0) 
        \prod_{t=0}^{T-1} \M^{\mu(t)}_{i(t)i(t+1)} \, \Pib(t+1),\\
  \text{where}\quad
  \Pib(t) &= \delta_{\w_{i(t)}\w(t)}
\end{aligned}
\end{equation}
%
We wish to maximise \eqref{eq:emq} subject to normalisation constraints.
Therefore we maximise the Lagrangian:
%
\begin{equation}\label{eq:lagrangesingle}
  \CL = Q(\M,\M') + \lambda\prn{1-\sum_i\pib'_i} + \sum_{\mu i} \lambda^\mu_i\prn{1-\sum_j\M'^\mu_{ij}}. 
\end{equation}
%
We find
%
\begin{equation}\label{eq:singlediffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \frac{N^\mu_{ij}}{\M'^\mu_{ij}} - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \frac{\alpha_i(0) \, \beta_i(0)}{\pib'_{i}} - \lambda,
\end{aligned}
\end{equation}
%
where the numerator $N^\mu_{ij}$ and the forward/backward variables $\alpha$ and $\beta$ are defined as:
%
\begin{equation}\label{eq:alphabeta}
\begin{aligned}
  N^\nu_{i(t)i(t+1)} &= \sum_{t=0}^{T-1} \pr\cond{i[t,t+1]),\w[0,T]}{\M} \\
     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, \alpha_i(t) \, \beta_j(t+1) \, \M^{\nu}_{i(t)i(t+1)} \, \Pib(t+1) ,\\
  \alpha_{i(t)}(t) &= \pr\cond{i(t),\w[0,t]}{\M} \\
     &= \sum_{i[0,t-1]} \pib_{i(0)} \, \delta_{\w_{i(0)}\w(0)} \prod_{s=0}^{t-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib(s+1), \\
  \beta_{i(t)}(t) &= \pr\cond{i(t),\w[t+1,T]}{\M} \\
     &= \sum_{i[t+1,T]} \prod_{s=t}^{T-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib(s+1).
\end{aligned}
\end{equation}
%
Note that, whilst disjoint sequences $\w[s,s']$ and $\w[t,t']$ are not independent, they are conditionally independent given the state at any time between the two intervals.
Therefore:
%
\begin{equation}\label{eq:stateprob}
  \pr\cond{i(t),\w[0,T]}{\M} = \alpha_{i(t)}(t) \, \beta_{i(t)}(t).
\end{equation}
%

These quantities usually suffer from underflow.
This can be avoided by using normalised forwar/backward variables:
%
\begin{equation}\label{eq:normalphabeta}
\begin{aligned}
  \tilde{\alpha}_i(t) &= \alpha_i(t) \prod_{s=0}^{t} \eta(s), \\
  \tilde{\beta}_i(t) &= \beta_i(t) \prod_{s=t+1}^{T} \eta(s),
\end{aligned}
\end{equation}
%
where $\eta[0,T]$ is chosen so that all of the $\tilde{\alpha}[0,T]$ are normalised: $\sum_i \tilde{\alpha}_i(t) = 1$.
These quantities have the interpretations:
%
\begin{equation}\label{eq:normintepr}
\begin{aligned}
  \tilde{\alpha}_{i(t)}(t) &= \pr\cond{i(t)}{\w[0,t],\M}, \\
  \eta(t) &= \pr\cond{\w(t)}{\w[0,t-1],\M}, \\
  \prod_{s=0}^t \eta(t) &= \pr\cond{\w[0,t]}{\M}.
\end{aligned}
\end{equation}
%
Note that the new backward variables $\tilde{\beta}$ are not normalised and have no particularly interesting interpretation.
However:
%
\begin{equation}\label{eq:normstateprob}
\begin{aligned}
  \widetilde{N}^\nu_{i(t)i(t+1)} &= \sum_{t=0}^{T-1} \pr\cond{i[t,t+1])}{\w[0,T],\M} \\
     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, \tilde{\alpha}_i(t) \, \tilde{\beta}_j(t+1) \, \eta(t+1) \, \M^{\nu}_{i(t)i(t+1)} \, \Pib(t+1) ,\\
  \pr\cond{i(t)}{\w[0,T],\M} &= \alpha_{i(t)}(t) \, \beta_{i(t)}(t).
\end{aligned}
\end{equation}
%
Then we can write
%
\begin{equation}\label{eq:singlediffsnorm}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \widetilde{N}^\mu_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \pib'_{i} } - \lambda,
\end{aligned}
\end{equation}
%
The factors of $\pr\cond{\w[0,T]}{\M}$ can be absorbed by a redefinition of the Lagrange multipliers.

Demanding a maximum and setting the derivatives to zero leaves us with the update rule:
%
\begin{equation}\label{eq:BWupdate}
\begin{aligned}
  \M'^\mu_{ij} &= \frac{ \widetilde{N}^\mu_{ij} }{ \sum_k \widetilde{N}^\mu_{ik} },\\
  \pib'_{i} &= \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \sum_j \tilde{\alpha}_j(0) \, \tilde{\beta}_j(0) }.
\end{aligned}
\end{equation}
%
This can be iterated until a maximum is found.

\subsection{Multiple sequences (Rabiner-Juang) \label{sec:rj}}

When we have multiple sequences of plasticity events, we will still want to maximise the Lagrangian \eqref{eq:lagrangesingle}, but now we will use the auxiliary function \eqref{eq:emqseveral}.
This results in
%
\begin{equation}\label{eq:multidiffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
    \frac{ N^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
    &= \pr\cond{\w^A[0,T]}{\M} \, \frac{ \sum_a \widetilde{N}^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
    \frac{ \alpha^a_i(0) \, \beta^a_i(0) }{ \pib'_{i} } - \lambda,\\
    &= \pr\cond{\w^A[0,T]}{\M} \,
    \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \pib'_{i} } - \lambda.
\end{aligned}
\end{equation}
%
Once again, we can absorb the factors of $\pr\cond{\w[0,T]}{\M}$ by a redefinition of the Lagrange multipliers.

Defining $\widetilde{N}^{A\mu}_{ij} = \sum_a \widetilde{N}^{a\mu}_{ij}$ and setting the derivatives to zero leaves us with the update rule:
%
\begin{equation}\label{eq:RJupdate}
\begin{aligned}
  \M'^\mu_{ij} &= \frac{ \widetilde{N}^{A\mu}_{ij} }{ \sum_k \widetilde{N}^{A\mu}_{ik} },\\
  \pib'_{i} &= \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \sum_{bj} \tilde{\alpha}^b_j(0) \, \tilde{\beta}^b_j(0) }.
\end{aligned}
\end{equation}
%
This can be iterated until a maximum is found.

\section{Sparseness promoting priors \label{sec:priors}}

When we have a prior distribution for the model, $\pr(\M)$, instead of maximising the likelihood \eqref{eq:like}, we can maximise the posterior:
%
\begin{equation}\label{eq:posterior}
  \pr\cond{\M}{\w[0,T]} = \frac{ \pr(\w[0,T],\M) }{ \pr(\w[0,T) } = \frac{ \pr\cond{\w[0,T]}{\M} \pr(\M) }{ \pr(\w[0,T) }.
\end{equation}
%
As the denominator is independent of the model, this is equivalent to maximising the numerator, \ie the joint distribution.

Using the same trick as before \eqref{eq:emq}, we can define an auxilliary function
%
\begin{equation}\label{eq:emqprior}
\begin{aligned}
  \widetilde{Q}(\M,\M') &= \sum_{i^A[0,T]} \pr\!\prn{i^A[0,T],\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M'} }{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M} } }\\
      &= \pr\!\prn{\M} Q(\M,\M') + \pr\!\prn{\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{\M'} }{ \pr\!\prn{\M} } }.
\end{aligned}
\end{equation}
%

We will consider priors of the form
%
\begin{equation}\label{eq:priorsum}
  \pr\!\prn{\M} = \frac{\exp\prn{-\beta\sum_{\mu ij}E^\mu_{ij}(\M^\mu_{ij})}}{Z(\beta)}.
\end{equation}
%
We will not put any prior on $\pib$, so its update rule will be unaffected.
The normalisation constant $Z(\beta)$ will not play any role either.
Now the derivatives read:
%
\begin{equation}\label{eq:multidiffs}
  \pdiff{\CL}{\M'^\mu_{ij}} 
    = \pr\!\prn{\w[0,T],\M} \prn{ \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - \pdiff{E^\mu_{ij}}{\M'^\mu_{ij}} }- \lambda^\mu_{i}.
\end{equation}
%
Once again, we can absorb the factors of $\pr\!\prn{\w[0,T],\M}$ by a redefinition of the Lagrange multipliers.

%\subsection*{Acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Appendices}
%\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{utcaps_sl}
\bibliography{maths,neuro,markov}

\end{document}
% ----------------------------------------------------------------